Notes on conformal manifolds

### 2025-05-27 Draft_MLFlatDirections.tex
update figures

In this work, we demonstrate the viability of this machine learning approach by applying it to a five-scalar subsector of a 13 scalar consistent truncation of three-dimensional chiral $\mathcal{N} = (2,0)$ supergravity arising from $AdS_3 \times S^3$ compactification from 6d supergravity, or from the compactification of type IIB supergravity on $AdS_3 \times S^3 \times T^4$. Our methodology combines gradient descent sampling of the flat direction manifold with a novel annealed importance sampling algorithm for polynomial symbolic regression. We show that this approach successfully identifies the underlying three-dimensional conformal manifold and automatically discovers the polynomial constraints that characterize it analytically.

@@

The scalar potential can be expressed in terms of the embedding tensor components as
%
{\setlength\arraycolsep{1.2pt}
	\begin{equation}	\label{eq: scalarpot_rephrased}
		\begin{aligned}
			V	&=	\frac1{12}\,\theta_{\bK\bL\bM\bN}\theta_{\bP\bQ\bR\bS}\Big(M^{\bK\bP}M^{\bL\bQ}M^{\bM\bR}M^{\bN\bS}-6\,M^{\bK\bP}M^{\bL\bQ}\eta^{\bM\bR}\eta^{\bN\bS}\\
			&\qquad\qquad\qquad\qquad\quad+8\,M^{\bK\bP}\eta^{\bL\bQ}\eta^{\bM\bR}\eta^{\bN\bS}-3\,\eta^{\bK\bP}\eta^{\bL\bQ}\eta^{\bM\bR}\eta^{\bN\bS}\Big)\\
			&\quad +\frac1{8}\,\theta_{\bK\bL}\theta_{\bP\bQ}\Big(2\,M^{\bK\bP}M^{\bL\bQ}-2\,\eta^{\bK\bP}\eta^{\bL\bQ}-M^{\bK\bL}M^{\bP\bQ}\Big)+4\,\theta\theta_{\bK\bL}M^{\bK\bL}-32\,\theta^2\,.
		\end{aligned}
	\end{equation}
}
In order to describe the $\mathcal{N} = (2,0)$ theory, it is useful to break the SO(8,4) group as

\begin{equation}	\label{eq: gl3gradingbar}
	\begin{aligned}
		\text{SO}(8,4)	&\longrightarrow	\enspace \text{GL}(3,\mathbb{R})\times\text{SO}(1,1)\times\text{SO}(4)_{\rm global}\,,	\\
		X^{\bM}		&\longrightarrow	\quad \{X^{\bar m},\; X_{\bar m},\; X^\bzero,\; X_\bzero,\; X^\balpha\}\,,
	\end{aligned}
\end{equation}
where $\bar m\in\llbracket1,3\rrbracket$ and $\balpha\in\llbracket9,12\rrbracket$ label the SL(3, $\mathbb{R}$) and SO(4)$_{\rm global}$ vector representations, respectively. In this basis, the SO(8,4)-invariant tensor reads
%
\begin{equation}	\label{eq: Paulieta}
	\eta_{\bM\bN}=
	\begin{pmatrix}
		0 & \delta_{\bar m}{}^{\bar n} & 0 & 0 &0\\
		\delta^{\bar m}{}_{\bar n} & 0 & 0 & 0 &0\\
		0 & 0 & 0 & 1 & 0\\
		0 & 0 & 1 & 0 & 0\\
		0 & 0 & 0 & 0 & -\delta_{\balpha\bbeta}
	\end{pmatrix}\,.
\end{equation}
The embedding tensor have the the following expression
\begin{equation}{\label{eq:SO84embeddings}}
		\theta_{\bar m \bar n \bar p\,\bzero}=-2\,\lambda\,\varepsilon_{\bar m \bar n \bar p}\,,		\qquad	\theta_{\bar m \bar n \bar p}{}^{\bzero}=\varepsilon_{\bar m \bar n \bar p}\,, \qquad	\theta_{\bar m \bar n}{}^{\bar p}{}_{\bzero}=\varepsilon_{\bar m \bar n \bar p}\,, 
\end{equation}
or equivalently

\begin{equation}
	X_{\bar m\bar n \bar p} = \varepsilon_{\bar m \bar n \bar p}\,,		\qquad
	X_{\bar m}{}^{\bar n \bar p} = \varepsilon_{\bar m \bar n \bar p}\,,	\qquad
	X^{\bar m}{}_{\bar n}{}^{\bar p} = \varepsilon_{\bar m \bar n \bar p}\,,	\qquad
	X^{\bar m\bar n}{}_{\bar p} = \varepsilon_{\bar m \bar n \bar p}\,,
\end{equation}
	%item Choice of truncation from 32 to 13 to 5 variables.
%\end{itemize}
Parametrizing the scalar matrix following \cite{Eloy:2021fhc}, we have for the scalar matrix 
\begin{equation}	\label{eq: scalarmatrix}
	M_{\bM\bN}
	=
	\begin{pmatrix}
		m+(\xi^2+\phi)m^{-1}(\xi^2-\phi)+2\xi^2	& 	(\xi^2+\phi)m^{-1}		&	0	&  0  &	-\sqrt2\,[1+(\xi^2+\phi)m^{-1}]\xi	\\
		m^{-1}(\xi^2-\phi)	& 	m^{-1}	&	0  	&  0  &	-\sqrt2\,m^{-1}\xi	\\
		0		&	0		&	e^{2\tilde\varphi}	&	0				&	0		\\
		0		&	0		&	0				&	e^{-2\tilde\varphi}	&	0		\\
		-\sqrt2\,\xi^T[1+m^{-1}(\xi^2-\phi)]	&	-\sqrt2\,\xi^Tm^{-1}	&	0	&	0	&	1+2\,\xi^Tm^{-1}\xi	\\
	\end{pmatrix}\,,
\end{equation}
which encodes 22 scalars of the theory ($22 = 32 - 10$, with 10 scalars gauge fixed using translations in the gauge group), which have been parametrized by 
$m = \nu\nu^{T}\in{\rm GL}(3,\mathbb{R})$ parametrizing the coset ${\rm GL}(3,\mathbb{R})/{\rm SO}(3)$, $\phi$ a $3\times3$ antisymmetric matrix,$\xi$ a $3\times4$ matrix, and $\xi^{2} = \xi \xi^{T}$, a dilaton $\tilde{\varphi}$. With this parametrization, the potential takes the from ref.~\cite{Eloy:2021fhc}:
\begin{equation} \label{eq:scalarpotential}
	\begin{aligned}
		V & = 4\,e^{-4\tilde\varphi}+2\,e^{-2\tilde\varphi}\Big[-\tr\left(m+m^{-1}\right)+\tr\left(\phi m^{-1}\phi\right) -2\,\tr\left(\phi m^{-1}\xi^{2}\right)-2\,\tr\left(\xi^{2}\right)\\
		&\qquad\quad-\tr\left(\xi^{2}m^{-1}\xi^{2}\right)  +\frac{1}{2}\,\det\left(m^{-1}\right)\left(1-\tr\left(\phi^{2}\right)-\tr\left(\xi^{4}\right)+\tr\left(\xi^{2}\right)^{2}\right) \\
		&\qquad\quad +\frac{1}{2}\,{\rm T}\left(m^{-1}(\xi^{2}-\phi),(\xi^{2}+\phi)m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\\
		&\qquad\quad +\frac{1}{4}\,{\rm T}\left(m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\Big], 
	\end{aligned}
\end{equation}
where ${\rm T}\left(A,B,C\right)=\varepsilon_{mnp}\,\varepsilon_{qrs}\,A^{mq}B^{nr}C^{ps}$. We can further restrict ourselves to a set of 13 scalars which is still a consistent truncation, with 
\begin{equation}
	\begin{aligned}
		\xi &= \begin{pmatrix}
					0 & 0 & 0 & x_{1} \\
					0 & 0 & 0 & x_{2} \\
					0 & 0 & 0 & x_{3}
				\end{pmatrix}, \\[5pt]
		\phi &= \begin{pmatrix}
					0 & x_{4} & x_{5} \\
					-x_{4} & 0 & x_{6} \\
					-x_{5} & -x_{6} & 0
				\end{pmatrix}, \\[5pt]
		\tilde{\varphi} &= \tilde{x}_{13}, \\[5pt]
		\nu &= e^{(6\,\tilde{x}_{7}+3\,\tilde{x}_{8}+\sqrt{3}\,\tilde{x}_{9})/6}
				\begin{pmatrix}
					1 & \frac{x_{10}}{\sqrt{2}} & \frac{x_{11}}{\sqrt{2}} + \frac{x_{10}x_{12}}{4} \\
					0 & e^{-\tilde{x}_{8}} & \frac{e^{-\tilde{x}_{8}}\,x_{12}}{\sqrt{2}} \\
					0 & 0 & e^{-(\tilde{x}_{8}+\sqrt{3}\,\tilde{x}_{9})/2}
				\end{pmatrix}.
	\end{aligned}
\end{equation}

@@

		\begin{figure}
			\centering
			\includegraphics[scale = 0.5]{Figures/Loss_grad_des.png}
			\caption{Evolution of the loss function during gradient descent and corresponding learning rate schedule, both plotted on a logarithmic scale. Dashed lines indicate epochs at which the optimizer was reinitialized.}
			\label{fig:loss_grad_des}
		\end{figure}

@@

		\begin{figure}[t]
			\centering
			\includegraphics[scale = 0.3]{Figures/triangular_plot_1_2_4_8_10_new.png} %names need to be changed.
			\caption{Triangular plot showing $2d$ projections and $1d$ histograms of the data after the gradient descent}
			\label{triangular_plot_1_2_4_8_10}
		\end{figure}

>>>>

In this work, we demonstrate the viability of this machine learning approach by applying it to a five-scalar subsector of a 13 scalar consistent truncation of three-dimensional chiral $\mathcal{N} = (1,1)$ supergravity arising from $AdS_3 \times S^3$ compactification from 6d supergravity, or from the compactification of type IIB supergravity on $AdS_3 \times S^3 \times T^4$. Our methodology combines gradient descent sampling of the flat direction manifold with a novel annealed importance sampling algorithm for polynomial symbolic regression. We show that this approach successfully identifies the underlying three-dimensional conformal manifold and automatically discovers the polynomial constraints that characterize it analytically.

@@

The scalar potential can be expressed in terms of the embedding tensor components as
%
{\setlength\arraycolsep{1.2pt}
	\begin{equation}	\label{eq: scalarpot_rephrased}
		\begin{aligned}
			V	&=	\frac1{12}\,\theta_{\bK\bL\bM\bN}\theta_{\bP\bQ\bR\bS}\Big(M^{\bK\bP}M^{\bL\bQ}M^{\bM\bR}M^{\bN\bS}-6\,M^{\bK\bP}M^{\bL\bQ}\eta^{\bM\bR}\eta^{\bN\bS}\\
			&\qquad\qquad\qquad\qquad\quad+8\,M^{\bK\bP}\eta^{\bL\bQ}\eta^{\bM\bR}\eta^{\bN\bS}-3\,\eta^{\bK\bP}\eta^{\bL\bQ}\eta^{\bM\bR}\eta^{\bN\bS}\Big)\\
			&\quad +\frac1{8}\,\theta_{\bK\bL}\theta_{\bP\bQ}\Big(2\,M^{\bK\bP}M^{\bL\bQ}-2\,\eta^{\bK\bP}\eta^{\bL\bQ}-M^{\bK\bL}M^{\bP\bQ}\Big)+4\,\theta\theta_{\bK\bL}M^{\bK\bL}-32\,\theta^2\,.
		\end{aligned}
	\end{equation}
}
\bd{redo}
Parametrizing the scalar matrix following \cite{Eloy:2021fhc}, we have for the scalar matrix 
\begin{equation}	\label{eq: scalarmatrix}
	M_{\bM\bN}
	=
	\begin{pmatrix}
		m+(\xi^2+\phi)m^{-1}(\xi^2-\phi)+2\xi^2	& 	(\xi^2+\phi)m^{-1}		&	0	&  0  &	-\sqrt2\,[1+(\xi^2+\phi)m^{-1}]\xi	\\
		m^{-1}(\xi^2-\phi)	& 	m^{-1}	&	0  	&  0  &	-\sqrt2\,m^{-1}\xi	\\
		0		&	0		&	e^{2\tilde\varphi}	&	0				&	0		\\
		0		&	0		&	0				&	e^{-2\tilde\varphi}	&	0		\\
		-\sqrt2\,\xi^T[1+m^{-1}(\xi^2-\phi)]	&	-\sqrt2\,\xi^Tm^{-1}	&	0	&	0	&	1+2\,\xi^Tm^{-1}\xi	\\
	\end{pmatrix}\,,
\end{equation}
which encodes 22 scalars of the theory ($22 = 32 - 10$, with 10 scalars gauge fixed using translations in the gauge group), which have been parametrized by 
$m = \nu\nu^{T}\in{\rm GL}(3,\mathbb{R})$ parametrizing the coset ${\rm GL}(3,\mathbb{R})/{\rm SO}(3)$, $\phi$ a $3\times3$ antisymmetric matrix,$\xi$ a $3\times4$ matrix, and $\xi^{2} = \xi \xi^{T}$, a dilaton $\tilde{\varphi}$. With this parametrization, the potential takes the from ref.~\cite{Eloy:2021fhc}:
\begin{equation} \label{eq:scalarpotential}
	\begin{aligned}
		V & = 4\,e^{-4\tilde\varphi}+2\,e^{-2\tilde\varphi}\Big[-\tr\left(m+m^{-1}\right)+\tr\left(\phi m^{-1}\phi\right) -2\,\tr\left(\phi m^{-1}\xi^{2}\right)-2\,\tr\left(\xi^{2}\right)\\
		&\qquad\quad-\tr\left(\xi^{2}m^{-1}\xi^{2}\right)  +\frac{1}{2}\,\det\left(m^{-1}\right)\left(1-\tr\left(\phi^{2}\right)-\tr\left(\xi^{4}\right)+\tr\left(\xi^{2}\right)^{2}\right) \\
		&\qquad\quad +\frac{1}{2}\,{\rm T}\left(m^{-1}(\xi^{2}-\phi),(\xi^{2}+\phi)m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\\
		&\qquad\quad +\frac{1}{4}\,{\rm T}\left(m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\Big], 
	\end{aligned}
\end{equation}
where ${\rm T}\left(A,B,C\right)=\varepsilon_{mnp}\,\varepsilon_{qrs}\,A^{mq}B^{nr}C^{ps}$. We can further restrict ourselves to a set of 13 scalars which is still a consistent truncation, with 
\begin{equation}
	\begin{aligned}
		\xi &= \begin{pmatrix}
					0 & 0 & 0 & x_{1} \\
					0 & 0 & 0 & x_{2} \\
					0 & 0 & 0 & x_{3}
				\end{pmatrix}, \\[5pt]
		\phi &= \begin{pmatrix}
					0 & x_{4} & x_{5} \\
					-x_{4} & 0 & x_{6} \\
					-x_{5} & -x_{6} & 0
				\end{pmatrix}, \\[5pt]
		\tilde{\varphi} &= \tilde{x}_{13}, \\[5pt]
		\nu &= e^{(6\,\tilde{x}_{7}+3\,\tilde{x}_{8}+\sqrt{3}\,\tilde{x}_{9})/6}
				\begin{pmatrix}
					1 & \frac{x_{10}}{\sqrt{2}} & \frac{x_{11}}{\sqrt{2}} + \frac{x_{10}x_{12}}{4} \\
					0 & e^{-\tilde{x}_{8}} & \frac{e^{-\tilde{x}_{8}}\,x_{12}}{\sqrt{2}} \\
					0 & 0 & e^{-(\tilde{x}_{8}+\sqrt{3}\,\tilde{x}_{9})/2}
				\end{pmatrix}.
	\end{aligned}
\end{equation}

We note here that all dilaton fields are denoted with a tilde. This convention is adopted in anticipation of a later redefinition of the form \( x_i = e^{\tilde{x}_i} \) for these fields. By reserving the notation without tildes for the exponentiated variables, we maintain a clear and consistent distinction for use in subsequent expressions.

@@

		\begin{figure}
			\centering
			\includegraphics[scale = 0.5]{Figures/Loss_grad_des.png}
			\caption{Evolution of the loss function during gradient descent and corresponding learning rate schedule, noted $\alpha$ on the left y-axis, both plotted on a logarithmic scale. Dashed lines indicate epochs at which the optimizer was reinitialized.}
			\label{fig:loss_grad_des}
		\end{figure}

@@

		\begin{figure}[t]
			\centering
			\includegraphics[scale = 0.3]{Figures/triangular_plot_1_2_4_8_10.png} %names need to be changed.
			\caption{Triangular plot showing $2d$ projections and $1d$ histograms of the data after the gradient descent}
			\label{triangular_plot_1_2_4_8_10}
		\end{figure}
